Reconstruction of Large Scale Structure by Nonlinear Displacement Fields

### 2016-09-25 ms_disp1/ms_disp1.tex
haoran

These algorithms include nonlinear Wiener filters in wavelet space [cite myself], distribution 
function transforms etc [cite Pengjie, etc.]. Standard BAO reconstruction algorithms smooth the 
nonlinear density field on linear scale ($\sim$10 Mpc$/h$) and reverse the large 
scale bulk flows by a negative Zel'dovich linear displacement [cite Eisenstein]. Here we propose a 
new reconstruction method that uses the nonlinear displacement field to recover the 
primordial density field. In the linear Lagrangian perturbation theory, the 
negative divergence of the the displacement field respect to Lagrangian coordinates 
gives the linear density field [cite Jeong]. Of course, the full displacement field is non-
observable, as it requires the initial distributions of matter, however there are 
many techniques to estimate the nonlinear displacement field from a final 
distribution of matter. For example, when a homogeneous initial matter distribution 
is assumed, there is a unique solution of curl-less displacement field to relate 
the initial and final distributions without shell-crossing. This solution can be 
solved by a metric transformation equation \citep{1995ApJS..100..269P,1998ApJS..115...19P}. In 1-dimensional case, the solution simplifies to a reordering of 
matter elements by Eulerian positions, and \cite{2016arXiv160907041Z} studies this 
case. The homogeneous and curl-less assumptions are generally valid in cosmology 
[cite or proof]. Shell-crossing does exist, and this effect can be quantified by 
$N$-body simulations.

@@

\begin{figure}[b] \centering
  \includegraphics[width=1.0\linewidth]{fig1.pdf}
  \caption{Dimensionless auto- and cross- power spectra between nonlinear density field $\delta$, linear density field $\delta_L$ and reconstructed density field $\delta_R$.}
  \label{fig.1}
\end{figure}
\begin{figure}[t] \centering
  \includegraphics[width=1.0\linewidth]{fig2.pdf}
  \caption{Correlation functions $r(\delta_L,\delta)$ and $r(\delta_L,\delta_R)$.}
  \label{fig.2}
\end{figure}

@@

\subsection{Reconstruction}\label{ss.reco}
In the simulation, we use particle-ID (PID) to record the initial (Lagrangian) location ${\bs 
q}$ of particles, and the information is tracked until the $z=0$ and we can get the 
Lagrangian displacement vector ${\bs \Psi}\equiv{\bs x}-{\bs q}$ for every 
particle. Then these vectors are interpolated onto the initial Lagrangian 
coordinates ${\bs q}$ of particles and we get the displacement field ${\bs \Psi}
({\bs q})$.
The raw reconstructed density field is given by the differential motion of matter 
elements,
\begin{equation}
    \delta_R=-\nabla\cdot{\bs \Psi}({\bs q}).
\end{equation}
Because the reconstruction processes are implemented on Lagrangian coordinates,
$\delta_R$ takes the coordinates of $\bs q$ instead of $\bs x$.
We just write $\bs q$'s Fourier wave number
$k_q$ as $k$ to simplify the expression.
In Fig.\ref{fig.1} we show the power spectrum of $\delta_R$ and its cross power with $\delta_L$. Despite of a lowered power of $\delta_R$ compared to $\delta$, it has a much higher cross power with $\delta_L$, up to a relatively smaller scale (higher $k$).

@@

\begin{figure}[t] \centering
  \includegraphics[width=1.0\linewidth]{fig3.pdf}
  \caption{Power spectra of the nonlinear density field and its
  filtered density field.}
  \label{fig.3}
\end{figure}

@@

\section{Discussion and conclusion}\label{sec.discussion}
Discussion goes here.

>>>>

These algorithms include nonlinear Wiener filters in wavelet space [cite myself], distribution 
function transforms etc [cite Pengjie, etc.]. Standard BAO reconstruction algorithms smooth the 
nonlinear density field on linear scale ($\sim$10 Mpc$/h$) and reverse the large 
scale bulk flows by a negative Zel'dovich linear displacement [cite Eisenstein]. Here we propose a 
new reconstruction method that uses the nonlinear displacement field to recover the 
primordial density field. In the linear Lagrangian perturbation theory, the 
negative divergence of the the displacement field respect to Lagrangian coordinates 
gives the linear density field [cite Jeong]. Of course, the full displacement field is non-
observable, as it requires the initial distributions of matter, however there are 
many techniques to estimate the nonlinear displacement field from a final 
distribution of matter. For example, when a homogeneous initial matter distribution 
is assumed, there is a unique solution of curl-less displacement field to relate 
the initial and final distributions without shell-crossing. This solution can be 
solved by a metric transformation equation \citep{1995ApJS..100..269P,1998ApJS..115...19P}. In 1-dimensional case, the solution simplifies to a reordering of 
matter elements by Eulerian positions, and \cite{2016arXiv160907041Z} studies this 
case. The homogeneous and curl-less assumptions are generally valid in cosmology 
[cite or proof]. Shell-crossing does exist, and this effect can be quantified by 
$N$-body simulations.
\begin{figure}[b] \centering
  \includegraphics[width=1.0\linewidth]{fig1.pdf}
  \caption{Dimensionless auto- and cross- power spectra between nonlinear density field $\delta$, linear density field $\delta_L$ and reconstructed density field $\delta_R$.}
  \label{fig.1}
\end{figure}
\begin{figure}[t] \centering
  \includegraphics[width=1.0\linewidth]{fig2.pdf}
  \caption{Correlation functions $r(\delta_L,\delta)$ and $r(\delta_L,\delta_R)$
  (solid lines). Their
  damping of the power spectra $W^2$ (dashed lines),
  and their best fit BAO damping models (dotted lines).}
  \label{fig.2}
\end{figure}

@@

@@

\subsection{Reconstruction}\label{ss.reco}
In the simulation, we use particle-ID (PID) to record the initial (Lagrangian) location ${\bs 
q}$ of particles, and the information is tracked until the $z=0$ and we can get the 
Lagrangian displacement vector ${\bs \Psi}\equiv{\bs x}-{\bs q}$ for every 
particle. Then these vectors are interpolated onto the initial Lagrangian 
coordinates ${\bs q}$ of particles and we get the displacement field ${\bs \Psi}
({\bs q})$.
The raw reconstructed density field is given by the differential motion of matter 
elements,
\begin{equation}
    \delta_R=-\nabla\cdot{\bs \Psi}({\bs q}).
\end{equation}
Because the reconstruction processes are implemented on Lagrangian coordinates,
$\delta_R$ takes the coordinates of $\bs q$ instead of $\bs x$.
We just write $\bs q$'s Fourier wave number
$k_q$ as $k$ to simplify the expression.
In Fig.\ref{fig.1} we show the power spectrum of $\delta_R$ and its cross power
with $\delta_L$. Despite of a lowered power of $\delta_R$ compared to $\delta$, it has a much higher 
cross power with $\delta_L$, up to a relatively smaller scale (higher $k$).
\begin{figure}[t] \centering
  \includegraphics[width=1.0\linewidth]{fig3.pdf}
  \caption{Power spectra of the nonlinear density field and its
  filtered density field.}
  \label{fig.3}
\end{figure}

@@

@@

Fig.\ref{fig.3} and Fig.\ref{fig.4} show their filtered power spectra.

\section{Discussion and conclusion}\label{sec.discussion}
We extract the actual displacement field of matter elements in cosmological $N$-body
simulations, and use this displacement field in Lagrangian coordinates to reconstruct
the primordial linear perturbations. The result shows a prominent improvement from
$r_{LN}$ to $r_{LR}$ in Fig.\ref{fig.2} -- recovering the lost linear information on
nearly an order of magnitude smaller scales. more tomorrow...
